### 2025-09-18 paper/sec/introduction.tex
Tweaks to intro.

Recent work on data provenance and data visualisation~\citep{psallidas18smoke,perera22,bond25} integrating
interactive provenance queries into charts and figures using \emph{transparent programming languages} which
track data provenance automatically as data flows through a computation. Visual outputs produced using these
platforms allow readers to interact with visual elements to see the relevant data. In most article charts and
figures usually only play a supporting role, however, with the main claims usually presented in natural
language. This language makes quantitative claims and refers to ``behind the scenes'' data analyses.

In principle transparent programming languages can also be used to create natural language which is bidirectly
linked to data, by computing key fragments of natural language from data. The following example, adapted from
\cite{moosavi21}, shows how the language presented in this paper can be used to achieve this:

>>>>

Recent work on data provenance and data visualisation~\citep{psallidas18smoke,perera22,bond25} proposed
integrating interactive ``provenance queries'' directly into charts and figures. These interactive queries
allow readers to interact with visual elements to see the data relevant to those elements. The key feature of
these approaches is that the interactions are provided automatically, via \emph{transparent programming
languages} (either query languages or general-purpose languages) which track dependencies as data flows
through a computation.

In most articles, however, charts and figures usually only play a supporting role. The main claims are usually
presented in natural language, language which also makes quantitative statements and refers to ``behind the
scenes'' data analyses. In principle transparent programming languages can also be used to create natural
language which is bidirectly linked to data, by computing key fragments of natural language from data. The
following example, adapted from \cite{moosavi21}, shows how the language presented in this paper can be used
to achieve this:
